---
title: "Paper: Zero-Adjusted Birnbaum-Saunders regression model"
author: "Vera Tomazella, Juvêncio S. Nobre, Gustavo, H.A. Pereira and Manoel Santos-Neto"
date: "August 14, 2018"
output: bookdown::html_document2
bibliography: bibliography.bib
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width=12,fig.height=8,echo=TRUE,warning=FALSE,message=FALSE)
```

# Application

## R packages
```{r}
library(gamlss.nl)
library(RBS) 
library(ggplot2) 
library(ssym)
```

## Data
In this application, the data set considered is the biaxial fatigue data analyzed by @rn:91, @lc:09, @Lemonte:2011aa, @vrc:12 and @Lemonte:2016aa. This data set has 46 observations and consists of lifes of a metal piece subjected to cyclic stretching and compressing, where the response variable $N$ denotes the number of cycles to failure of the metal specimen and the explanatory variable $W$ is the work per cycle ($mj/m^3$).  Here, we use the response variable on a reduced scale to facilitate the estimation procedure, i.e., $Y \times 100 \equiv N$.
```{r}
data("Biaxial", package="ssym")
N <- Biaxial$Life
N <- N/100;N
W <- Biaxial$Work;W
```


## RBS nonlinear regression model

 Therefore, we fit a RBS nonlinear regression model
\[
\mu_i = \beta_1 \textrm{e}^{\frac{\beta_2}{w_i}}, \quad i=1,\ldots,46,
\]
where $Y_i \sim \text{RBS}(\mu_i,\sigma)$. The adjustment was performed using the function **nlgamlss()** of the package **gamlss.nl** whose objective  is to allow nonlinear fitting within a GAMLSS model. The estimates and standard errors obtained were
```{r}
mod1 <- nlgamlss(y=N,mu.fo=~p1*exp(p2/W),sigma.formula = ~1,mu.start =c(1,40),sigma.start =1,family = RBS(mu.link = "identity",sigma.link = "identity"))
mod1$converged
summary(mod1)
```


```{r fig,include=FALSE}
envelope.nlrbs <- function(model,k=19, color = "grey50", xlabel = "Theorical Quantile",ylabel = "Empirical Quantile",font="serif")
{
 
  n <- model$N
  td  <- model$residuals
  sigma <- model$sigma.fv[1]
  mu <- model$mu.fv
  re <- matrix(0,n,k)

  for(i in 1:k)
  { #inicio do for do residuos
    y1 <- mapply(rRBS,n=1,mu=mu,sigma=sigma)
    model1 <- nlgamlss(y=y1,mu.fo=model$mu.formula,sigma.formula = ~1,mu.start =c(1.276,40.954),sigma.start =9.817,family = RBS(mu.link = "identity",sigma.link = "identity"))
    rdf=model1$residuals
    re[,i]=sort(rdf)
  }
  e10 <- numeric(n)
  e20 <- numeric(n)
  e11 <- numeric(n)
  e21 <- numeric(n)
  e12 <- numeric(n)
  e22 <- numeric(n)
  for(l in 1:n)
  {#inicio do Bootstrap
    eo = sort(re[l,])
    e10[l] = eo[ceiling(k*0.01)]
    e20[l] = eo[ceiling(k*(1 - 0.01))]
    e11[l] = eo[ceiling(k*0.05)]
    e21[l] = eo[ceiling(k*(1 - 0.05))]
    e12[l] = eo[ceiling(k*0.1)]
    e22[l] = eo[ceiling(k*(1 - 0.1))]
  }#fim do Bootstrap
  
  a <- qqnorm(e10, plot.it = FALSE)$x
  r <- qqnorm(td, plot.it = FALSE)$x
  xb = apply(re, 1, mean)
  rxb <- qqnorm(xb, plot.it = FALSE)$x
  
  df <- data.frame(r=r,xab=a,emin=cbind(e10,e11,e12),emax=cbind(e20,e21,e22),xb=xb,td=td,rxb=rxb)
  ggplot(df,aes(r,td))+geom_ribbon(aes(x=xab, ymin=emin.e10, ymax=emax.e20),fill=color,alpha=0.5)  + geom_ribbon(aes(x=xab, ymin=emin.e11, ymax=emax.e21),fill=color,alpha=0.5) + geom_ribbon(aes(x=xab, ymin=emin.e12, ymax=emax.e22),fill=color,alpha=0.5) +scale_fill_gradient(low = "grey25", high = "grey75")+ geom_point() + geom_line(aes(rxb,xb),lty=2)+xlab(xlabel)+ylab(ylabel)+theme(text=element_text(size=30,family=font)) 
}
```
Figure \@ref(fig:fig1) shows the QQ-plot with simulated envelope of quantile residuals and the scatter plot with the fitted model.  Figures \@ref(fig:fig1)(a) and \@ref(fig:fig2) are used to verify the distribution assumption of the model and they do not show unusual features. Note that, in Figure \@ref(fig:fig1)(b), the fitted line (proposed model) has a good agreement with the real data.

```{r fig1,fig.cap="QQ-plot with simulated envelope of quantile residuals (a) and scatter-plot with fitted model (b).",fig.show = 'hold',out.width = '50%'}
set.seed(13171619)
envelope.nlrbs(mod1,k=100)
df <- data.frame(N=N,W=W,fv=mod1$mu.fv)
ggplot(df,aes(W,N)) + geom_point(size=5) + geom_line(aes(W,fv),col=2)+ xlab("Work per cycle") + ylab("Number of cycles to failure (x 100)")+ theme(text=element_text(size=30,family="serif"))
```

```{r fig2, fig.show = 'hold',out.width = '50%',fig.cap="Residuals plots."}
df<- data.frame(res=mod1$residuals,ind=1:mod1$N,fv=mod1$mu.fv)
ggplot(df,aes(ind,res)) + geom_point(size=5) + geom_hline(yintercept = c(-3,3),lty=2) + xlab("Index") + ylab("Quantile Residual") + theme(text=element_text(size=30,family="serif"))
ggplot(df,aes(fv,res)) + geom_point(size=5) + geom_hline(yintercept = c(-3,3),lty=2)+ xlab("Fitted Values") + ylab("Quantile Residual") + theme(text=element_text(size=30,family="serif"))

```

```{r include=FALSE}
model <- mod1
x1 <- exp(47.957/W)
x2 <- (1.276/W)*exp(47.957/W)
x <- cbind(x1,x2)
z <- matrix(1,nrow=model$N)
y <- model$y
p <- ncol(x)
q <- ncol(z)
linkstr <- "identity"
linkobj <- make.link(linkstr)
linkfun <- linkobj$linkfun
linkinv <- linkobj$linkinv
mu.eta <- linkobj$mu.eta
sigma_linkstr <- "identity"
sigma_linkobj <- make.link(sigma_linkstr)
sigma_linkfun <- sigma_linkobj$linkfun
sigma_linkinv <- sigma_linkobj$linkinv
sigma_mu.eta <- sigma_linkobj$mu.eta
B = function(Delta, I, M) {
  B = (t(Delta) %*% (I - M) %*% Delta)
  return(B)
}
loglik <- function(vP) {
  betab = vP[1:p]
  alpha = vP[-(1:p)]
  eta = as.vector(x %*% betab)
  tau = as.vector(z %*% alpha)
  mu = linkinv(eta)
  sigma = sigma_linkinv(tau)
  f <- 0.5 * sigma - 0.5 * log((sigma + 1)) - 0.5 * log(mu) - 
    1.5 * log(y) + log((sigma * y) + y + (sigma * mu)) - 
    y * (sigma + 1)/(4 * mu) - (sigma * sigma * mu)/(4 * 
                                                       y * (sigma + 1)) - 0.5 * log(16 * pi)
  return(sum(f))
}
muest <- model$mu.coefficients
sigmaest <- model$sigma.coefficients
x0 <- c(muest, sigmaest)
h0 <- hessian(loglik, x0)
Ldelta = h0[(p + 1):(p + q), (p + 1):(p + q)]
Lbeta = h0[1:p, 1:p]
b11 = cbind(matrix(0, p, p), matrix(0, p, q))
b12 = cbind(matrix(0, q, p), solve(Ldelta))
B1 = rbind(b11, b12)
b211 = cbind(solve(Lbeta), matrix(0, p, q))
b212 = cbind(matrix(0, q, p), matrix(0, q, q))
B2 = rbind(b211, b212)
b311 = cbind(matrix(0, p, p), matrix(0, p, q))
b312 = cbind(matrix(0, q, p), matrix(0, q, q))
B3 = rbind(b311, b312)
mu <- model$mu.fv
sigma <- model$sigma.fv
eta <- linkfun(mu)
ai <- mu.eta(eta)
dmu <- (-1/(2 * mu)) + sigma/((y * sigma) + y + (sigma*mu)) + ((sigma + 1) * y)/(4 * (mu^2)) - (sigma^2)/(4 * 
                                                                                                          y * (sigma + 1))
  Deltamu <- crossprod(x, diag(ai * dmu))
  tau <- sigma_linkfun(sigma)
  bi <- sigma_mu.eta(tau)
  dsigma <- (y + mu)/((sigma * y) + y + (sigma * mu)) - 
    y/(4 * mu) - (sigma * (sigma + 2) * mu)/(4 * (sigma + 
                                                    1) * (sigma + 1) * y) + sigma/(2 * (sigma + 1))
  Deltasigma <- crossprod(z, diag(bi * dsigma))
  Delta <- rbind(Deltamu, Deltasigma)
  BT <- B(Delta, solve(h0), B3)
  autovmaxthetaPC <- eigen(BT, symmetric = TRUE)$val[1]
  vetorpcthetaPC <- eigen(BT, symmetric = TRUE)$vec[, 1]
  dmaxG.theta <- abs(vetorpcthetaPC)
  vCithetaPC <- 2 * abs(diag(BT))
  Cb0 <- vCithetaPC
  Cb.theta <- Cb0/sum(Cb0)
  BM <- B(Delta, solve(h0), B1)
  autovmaxbetaPC <- eigen(BM, symmetric = TRUE)$val[1]
  vetorpcbetaPC <- eigen(BM, symmetric = TRUE)$vec[, 1]
  dmaxG.beta <- abs(vetorpcbetaPC)
  vCibetaPC <- 2 * abs(diag(BM))
  Cb1 <- vCibetaPC
  Cb.beta <- Cb1/sum(Cb1)
  BD <- B(Delta, solve(h0), B2)
  autovmaxdeltaPC <- eigen(BD, symmetric = TRUE)$val[1]
  vetordeltaPC <- eigen(BD, symmetric = TRUE)$vec[, 1]
  dmaxG.alpha = abs(vetordeltaPC)
  vCideltaPC = 2 * abs(diag(BD))
  Cb2 = vCideltaPC
  Cb.alpha = Cb2/sum(Cb2)
  local.infl <- list(dmax.beta = dmaxG.beta, dmax.alpha = dmaxG.alpha, 
                 dmax.theta = dmaxG.theta, Ci.beta = Cb.beta, Ci.alpha = Cb.alpha, 
                 Ci.theta = Cb.theta)
 
```

Infuence diagnostics for the nonlinear BS regression model are presented in Figure \@ref(fig:fig3). Figure \@ref(fig:fig3) displays index plots of $\mathrm{C}_i$ and $|\mathbf{d}_\text{max}|$, from where observation \#1 is detected as potentially influential. Changes are not detected when the observation is removed.

```{r fig3,fig.show = 'hold',out.width = '33%',fig.cap="Index plots of C_i and d_{max}."}  
  Cibeta <- local.infl$Ci.beta
  Cialpha <- local.infl$Ci.alpha
  Citheta <- local.infl$Ci.theta
  
  dmbeta <- local.infl$dmax.beta
  dmalpha <- local.infl$dmax.alpha
  dmtheta <- local.infl$dmax.theta 
  
  index <- 1:length(Cibeta)
  CbxI <- data.frame(index,Cibeta,Cialpha,Citheta)
  CbxI. <- CbxI[c(1),]
  ggplot(CbxI,aes(index,Cibeta)) + geom_point(size=10)+ ylim(0,1) + geom_text(data=CbxI.,aes(index,Cibeta,label=index),hjust=2,size=10) + xlab("Index") + ylab(expression(C[i]))+ theme(text=element_text(size=50,family="serif"))
  ggplot(CbxI,aes(index,Cialpha)) + geom_point(size=10)+ ylim(0,1) + geom_text(data=CbxI.,aes(index,Cialpha,label=index),hjust=2,size=10) + xlab("Index") + ylab(expression(C[i]))+ theme(text=element_text(size=50,family="serif"))
  ggplot(CbxI,aes(index,Citheta)) + geom_point(size=10)+ ylim(0,1) + geom_text(data=CbxI.,aes(index,Citheta,label=index),hjust=2,size=10) + xlab("Index") + ylab(expression(C[i]))+ theme(text=element_text(size=50,family="serif"))

  I <- 1:length(dmbeta)
  dxI <- data.frame(I,dmbeta,dmalpha,dmtheta)
  dxI... <- dxI[c(1),]
  ggplot(CbxI,aes(I,dmbeta)) + geom_point(size=10)+ ylim(0,1)  + xlab("Index") + ylab(expression(d[max])) +  theme(text=element_text(size=50,family="serif"))
  ggplot(CbxI,aes(I,dmalpha)) + geom_point(size=10)+ ylim(0,1) + geom_text(data=dxI...,aes(I,dmalpha,label=I),hjust=2,size=10) + xlab("Index") + ylab(expression(d[max]))+ theme(text=element_text(size=50,family="serif"))
  ggplot(CbxI,aes(I,dmtheta)) + geom_point(size=10)+ ylim(0,1) + xlab("Index") + ylab(expression(d[max]))+ theme(text=element_text(size=50,family="serif"))
```

Finally, we have that the mean of the number of cycles ($\times 100$) to failure of the metal specimen can be described by $\hat \mu(w_i) = 1.276\times \textrm{e}^{\frac{47.957}{w_i}}$.

## References